---
title: "directMarketingArules"
author: "NareshVemula"
date: "September 16, 2016"
output: html_document
---
exploring direct marketing dataset to see for patterns 
```{r}
directMarket=read.csv("../../dirmktg.csv")
str(directMarket)
table(directMarket$outcome)
```
preparing the data 
binning the variable earnings as factor variable
```{r}
attach(directMarket)
directMarket$incomeBins=cut(earnings,breaks = c(5000,85000,110000,150000),
labels=c("below 85000","below 110000","below 150000"),
include.lowest = T)
table(directMarket$incomeBins)
directMarket$zip.code=as.factor(zip.code)
```
taking the necessary features by removing name and zipcode
-------------------------------------------------------------
binning age variable into three categories
```{r}
directMarket$zip.code=NULL

directMarket$earnings=NULL
directMarket$name=NULL  
str(directMarket)
quantile(age)
directMarket$ageBins=cut(age,breaks = c(quantile(age)),labels=c("below 28","below 42","below 56","below 69"),include.lowest = T)
table(directMarket$ageBins)
directMarket$age=NULL
table(directMarket$outcome)
```
loading the packages arules and arulesViz
```{r}
library(arules)
library(arulesViz)
```
computing lift for response variable considering independent variable
```{r}
directMarketArules=apriori(directMarket,parameter =list(confidence=0.7,support=0.05),appearance = list(default="lhs",rhs=c("outcome=response","outcome=no response")),control = list(verbose=F))
directMarketArules=sort(directMarketArules,decreasing = T,by="lift")
inspect(directMarketArules[1:5])
```
removing duplicate association rules
```{r}
directMarketArulesRedundant=is.subset(directMarketArules,directMarketArules)
directMarketArulesRedundant[lower.tri(directMarketArulesRedundant,diag = T)]=NA
directMarketArulesRedundant=colSums(directMarketArulesRedundant,na.rm = T)>=1
which(directMarketArulesRedundant)
#removing redundant data
directMarketArules=directMarketArules[!directMarketArulesRedundant]
directMarketArules=sort(directMarketArules,decreasing = T,by="lift")
inspect(directMarketArules[1:5])
```
plotting the associations
```{r}
library(iplots)
directMarketArules
plot(directMarketArules,interactive = F,method="graph",measure = "lift")
#plot(directMarketArules,interactive = T,method="iplots",measure = "lift")

```