---
title: "LDA and Logistic Regression"
author: "Naresh"
date: "September 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

working with stock market data
```{R}
library(ISLR)
str(Smarket)
?Smarket
summary(Smarket)
cor(Smarket[,-9])
pairs(Smarket)
attach(Smarket)
plot(Volume)
```

working with logistic regression
```{r}
smarket1=glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,data = Smarket,family = binomial)
summary(smarket1)
coef(smarket1)
summary(smarket1)$coef[,4]
obs=predict(smarket1,data=smarket1,type = 'response')
contrasts(Direction)
table(TrainProbs>.5,Smarket$Direction)[1][1]
measures=rep(1:4)
accuracy=function(x)
{
  return ((x[1,1]+x[2,2])/sum(x))
}
acc=accuracy(table(TrainProbs>.5,Smarket$Direction))
acc
```
train and test sets
```{r}
train=(Smarket$Year<2005)
SmarketTest=Smarket[!train,]
fitTrain=glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,data = Smarket,
             subset = train,family = binomial)
testProbs=predict(fitTrain,newdata = SmarketTest,type='response')
testAcc=accuracy(table(testProbs>0.5,SmarketTest$Direction))
testAcc
```
improved logit model
```{r}
fitTrain=glm(Direction~Lag1+Lag2,data = Smarket,
             subset = train,family = binomial)
testProbs=predict(fitTrain,newdata = SmarketTest,type='response')
testAcc=accuracy(table(testProbs>0.5,SmarketTest$Direction))
testAcc
```
LDA---Linear Discriminant Analysis
---------------------------------
```{r}
library(MASS)
lda1=lda(Direction~Lag1+Lag2,data = Smarket,subset = train)
lda1
op=par(mar=rep(2,4))
plot(lda1)
par(op)
predictTest=predict(lda1,SmarketTest)
names(predictTest)
ldaacc=accuracy(table(predictTest$class,SmarketTest$Direction))
ldaacc

ldaacc.5=accuracy(table(predictTest$posterior[,1]>.46,SmarketTest$Direction))
ldaacc.5

```

---------------QDA-----

```{r}
qda1=qda(Direction~Lag1+Lag2,data = Smarket,subset = train)
qda1
predictTest=predict(qda1,SmarketTest)
names(predictTest)
qdaacc=accuracy(table(predictTest$class,SmarketTest$Direction))
qdaacc
```
------------Working with KNN------
requirements
1. matrix of predictors
2.matrix of test set predictors
3.vector of dependent varaible belongs to train
4. K value
```{r}
library(class)
train.x=as.matrix(Smarket[!train,c('Lag1','Lag2')])
train.y=as.matrix(Smarket[!train,c('Direction')])
test.x=as.matrix(Smarket[train,c('Lag1','Lag2')])
set.seed(1)
knn1=knn(train = train.x,cl = train.y,test = test.x,k=1)
knn1
knnacc=accuracy(table(knn1,Smarket[train,'Direction']))
knnacc
```

increasing k value to 3
-----------------------
```{r}

knn1=knn(train = train.x,cl = train.y,test = test.x,k=3)
knn1
knnacc=accuracy(table(knn1,Smarket[train,'Direction']))
knnacc
```

---applying KNN to caravan data----
```{r}
dim(Caravan)
summary(Caravan$Purchase)
348/5474
```

---large scale difference between units will affect the performance of KnN, we will solve this by standardizing the predictors.
```{r}
StatndardX=scale(Caravan[,-86])
```
splitting data into test and train
```{r}
test=1:1000
trainX=StatndardX[-test,]
testX=StatndardX[test,]
trainY=Caravan[-test,'Purchase']
testY=Caravan[test,'Purchase']
```

building the model, k=1
```{r}
set.seed(1)
knnCaravan=knn(train = trainX,cl = trainY,test = testX,k=1)
knnacc=accuracy(table(knnCaravan,testY))
knnacc
tp=9/(9+68)
tp
```
building the model, k=3
```{r}
knnCaravan=knn(train = trainX,cl = trainY,test = testX,k=3)
knnacc=accuracy(table(knnCaravan,testY))
knnacc
tp=5/(5+21)
tp
```

ISLr exercises
using prior probability of down as threshold
```{r}
?Weekly
attach(Weekly)
summary(Weekly)
pairs(Weekly[,-9])
Q10=glm(Direction~Year+Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=Weekly,family  = binomial)
summary(Q10)
predictWeekly=predict(Q10,Weekly)
table(predictWeekly>0.5,Weekly$Direction)
WeeklyAcc=accuracy(table(Weekly$Direction,predictWeekly>0.444))
WeeklyAcc
table(Weekly$Direction)
```
trainng and test
```{r}
library(caTools)
trainSplit=Weekly$Year<2009
train=Weekly[trainSplit,]
test=Weekly[!trainSplit,]
Q10.1=glm(Direction~Year+Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=train,family  = binomial)
summary(Q10)
predictWeekly=predict(Q10,newdata=test)
table(test$Direction,predictWeekly>0.5)
WeeklyAcc=accuracy(table(test$Direction,predictWeekly>0.444))
WeeklyAcc
table(test$Direction)
```
LDA
```{r}
library(MASS)
LDAWeekly=lda(Direction~Year+Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=train)
LDAWeekly
predictWeekly=predict(LDAWeekly,newdata=test)
table(test$Direction,predictWeekly$class)
WeeklyAcc=accuracy(table(test$Direction,predictWeekly$class))
WeeklyAcc
```