---
title: "NonlinearTechniques"
author: "NareshVemula"
date: "September 10, 2016"
output: html_document
---
Non linear models
-----------------

```{r}
library(plotly)
require(ISLR)
attach(Wage)
```

polynomial regression
poly function uses orthogonal polynomials
```{r}
wagePolyReg=lm(wage~poly(age,4),data=Wage)
summary(wagePolyReg)
```

plotting the above function using SE's
lwd---line width
matlinies--to plot two or more lines at a time
lyt--line type, 2 is a broken line
uncorrelated polynomials/orthogonal polynomials
```{r}
agelims=range(age)
age.grid=seq(from=agelims[1],to=agelims[2] )
predictAgegrid=predict(wagePolyReg,newdata=list(age=age.grid),se=T)
se.bands=cbind(predictAgegrid$fit+2*predictAgegrid$se,predictAgegrid$fit-2*predictAgegrid$se)
plot(age,wage,col='darkgrey')
lines(age.grid,predictAgegrid$fit,lwd=2,col="blue")
matlines(age.grid,se.bands,col='blue',lyt=2)
```

polynomial using normal math

```{r}
wageMathPolyReg=lm(wage~age+I(age^2)+I(age^3)+I(age^4),data=Wage)
summary(wageMathPolyReg)
```

```{r}
plot(fitted(wagePolyReg),fitted(wageMathPolyReg))
```
this work with Linear Regression, ANOVA()
```{r}
fitEdu=lm(wage~education,data=Wage)
fitEduAge=lm(wage~education+age,data=Wage)
fitEduAgesq=lm(wage~education+poly(age,2),data=Wage)
fitEduAgecb=lm(wage~education+poly(age,3),data=Wage)
anova(fitEdu,fitEduAge,fitEduAgesq,fitEduAgecb)
```
Polynomial logistic regression
```{r}
plot(age,wage,col='darkgrey')
wagePolyglm=glm((wage>250)~poly(age,3),data=Wage,family=binomial)
summary(wagePolyglm)
predictAge=predict(wagePolyglm,newdata = list(age=age.grid),se=T)
se.bands=predictAge$fit+cbind(wagePolyglm=0,lower=-2*predictAge$se.fit,upper=2*predictAge$se.fit)
se.bands[1:5,]
```
as I have domputeted the odds on logit scale, I am converting it to the probability scale using logistic regression exp.
$$p=\frac{e^\eta}{1+e^\eta}.$$
```{r}
prob.bands=exp(se.bands)/(1+exp(se.bands))
matplot(age.grid,prob.bands,col='blue',lwd=c(2,1,1),lty=c(1,2,2),
        type="l",ylim=c(0,.1))
#plotting 1's and 0's on the plot
points(jitter(age),I(wage>250)/10,pch='l',cex=.5)
```

splines
more flexible and offers wider scope of possibilites by regressing differently in different area's using knots.
bs---cubic splines
```{r}
require(splines)
fit=lm(wage~bs(age,knots = c(25,40,60)),data=Wage)
plot(age,wage,col='darkgrey')
lines(age.grid,predict(fit,list(age=age.grid)),col='darkgreen',lwd=2)
summary(fit)
#adding knots
abline(v=c(25,40,60),lty=2,col='darkgreen')
```
fiiting a smoothing spline with DOF
```{r}
fit=smooth.spline(age,wage,df=16)
lines(fit,col="red",lwd=2)
```

Using Leave on out cv
to pick the tuning parameter lamdba or DOF
```{r}
fit=smooth.spline(age,wage,cv = T)
fit
lines(fit,col='black',lwd=2)
```

GAM's
s=smoothing spline
```{r fig.width=10, fig.height=5}
require(gam)
gam=gam(wage~s(age,df=4)+s(year,df=4)+education,data=Wage)
par(mfrow=c(1,3))
plot(gam,se=T)
plot.gam(gam,se=T)
```
gam for buinary response
```{r}
gam=gam(I(wage>250)~s(age,df=4)+s(year,df=4)+education,data=Wage)
plot(gam)
```
testing the necessity of non lineaity in year
```{r}
gamNLTest=gam(I(wage>250)~s(age,df=4)+year+education,data=Wage)
plot(gamNLTest)
anova(gamNLTest,gam,test='Chisq')
```
gam can be used to plot lm and glm's
```{r}
lm1=lm(wage~ns(age,df=4)+ns(year,df=4)+education,data=Wage)
plot.gam(lm1,se=T)
?gam
```
working on exercise
```{r}
plot(x,y)
exLM=lm(y~x)
summary(exLM)
exNLM=lm(y~1+poly(x,2))
exNLM=lm(y~1+x+I(x^2))

summary(exNLM)
plot.gam(exNLM)
```



