---
title: "PolynomialsAndSplines"
author: "NareshVemula"
date: "September 13, 2016"
output: html_document
---
Loading the necessary packages
will work with wage data to implement Polynomial reg, Splines,smoothed Splines and GAM's
```{r}
library(ISLR)
attach(Wage)
```
Polynomial Regression and Step Functions
Fitting only one variable in regression method with polynomials
poly functions helps us to reduce the collinearity by fitting orthogonal polynomials
which are linear combinations of all polynomials.
```{r}
fitPolyWage=lm(wage~poly(age,4),data = Wage)
coef(summary(fitPolyWage))
```
adding raw=TRUE argument to the poly function
this gives us coefficients which are confusing rather than revealing
```{r}
fitPolyWage=lm(wage~poly(age,4,raw=F),data = Wage)
coef(fitPolyWage)
```
predicting wage using machine generated data
computing standard errors to plot
```{r}
ageRange=range(age)
ageTest=seq(from=ageRange[1],to=ageRange[2])
predictAge=predict(fitPolyWage,newdata = list(age=ageTest),se.fit =T)
seBands=cbind(predictAge$fit+2*predictAge$se.fit,predictAge$fit-2*predictAge$se.fit)
```
plotting the above predictions and SE's
```{r}
par(mfrow=c(1,2),mar=c(4.5,4.5,1,1),oma=c(0,0,4,0))
plot(age,wage,xlim = ageRange,ylim = range(wage),cex=.5,col="darkgrey")
lines(ageTest,predictAge$fit,lwd=2,col="blue")
matlines(ageTest,seBands,lwd=1,col="red",lty=3)
```
Using Anove to Select the optimum model based on RSS and DF 
```{r}
fitPolyWage1=lm(wage~age,data = Wage)
fitPolyWage2=lm(wage~poly(age,2),data = Wage)
fitPolyWage3=lm(wage~poly(age,3),data = Wage)
fitPolyWage4=lm(wage~poly(age,4),data = Wage)
fitPolyWage5=lm(wage~poly(age,5),data = Wage)

anova(fitPolyWage1,fitPolyWage2,fitPolyWage3,fitPolyWage4,fitPolyWage5)
coef(summary(fitPolyWage5))
```
classifying people with wage greater than 250k using CV
```{r}
fitGlm=glm(I(wage>250)~poly(age,4),family  = "binomial",data = Wage)
predictAge=predict(fitGlm,newdata = list(age=ageTest),se.fit = T)
```
computing CI using logistic function
```{r}
modelFit=exp(predictAge$fit)/(1+exp(predictAge$fit))
seGlm=cbind(predictAge$fit+2*predictAge$se.fit,predictAge$fit-2*predictAge$se.fit)
#changing SE to logistic function scale
seGlm=exp(seGlm)/(1+exp(seGlm))
```
plotting se vs preds
```{r}
plot(y=I(wage>250),x=age,xlim = ageRange,ylim=c(0,0.2),col="darkgrey")
points(jitter(age),I((wage>250)/5),cex=.5,pch="|")
lines(ageTest,modelFit,col="blue")
matlines(ageTest,seGlm,lwd="1",col="blue",lty=2)
```
fitting step functions using cut in r
```{r}
table(cut(age,4))
fitStep=lm(wage~cut(age,4),data = Wage)
summary(fitStep)
```
fitting splines
```{r}
library(splines)
fitLmSplines=lm(wage~bs(age,knots = c(25 ,40 ,60)),data = Wage)
summary(fitLmSplines)
predictWageLs=predict(fitLmSplines,newdata = list(age=ageTest),se.fit = T)
plot(age,wage,col="dark grey")
lines(ageTest,predictWageLs$fit,lwd=2)
lines(ageTest,predictWageLs$fit+2*predictWageLs$se.fit,lty='dashed',col="red",lwd=2)
lines(ageTest,predictWageLs$fit-2*predictWageLs$se.fit,lty='dashed',col="red",lwd=2)

```
dimensions of a cubic spline with three knots is seven (intercept + six basis functions)
```{r}
dim(bs(age,knots=c(25,20,60)))
dim(bs(age,df=6))
attr(bs(age,df=6),"knots")
```
fitting natural splines
```{r}
fitNsSplines=lm(wage~ns(age,knots = c(25 ,40 ,60)),data = Wage)
summary(fitNsSplines)
predictWage=predict(fitNsSplines,newdata = list(age=ageTest),se.fit = T)
plot(age,wage,col="dark grey")
lines(ageTest,predictWage$fit,lwd=2)
lines(ageTest,predictWage$fit+2*predictWage$se.fit,lty='dashed',col="blue",lwd=2)
lines(ageTest,predictWage$fit-2*predictWage$se.fit,lty='dashed',col="blue",lwd=2)
```
working with smoothing splines
```{r}
plot(age,wage,xlim = ageRange,cex=.5,col="darkgrey")
title("smoothing splines")
fitSmooth=smooth.spline(age,wage,df=16)
fitSmoothCV=smooth.spline(age,wage,cv = T)
fitSmoothCV
fitSmoothCV$df
lines(fitSmooth,col="red",lwd=2,lty="dashed")
lines(fitSmoothCV,col="blue",lwd=2)
legend("topright",legend = c("16 Df","6 DF"),col=c("red","blue"),lty=1,lwd=2,cex=.8)
predictSmooth=predict(fitSmoothCV,newdata=list(age=ageTest),se.fit=T)
lines(predictSmooth$x,predictSmooth$y)
```
using local regression
we can also use locfit library for local regression
```{r}
plot(age,wage,cex=.5,xlim = ageRange,col="darkgrey")
title("local Regression")
fitLocalReg=loess(wage~age,span = 0.2,data = Wage)
fitLocalReg50=loess(wage~age,span = 0.5,data = Wage)
predictLoess=predict(fitLocalReg,newdata =data.frame(age=ageTest),se=T)
predictLoess50=predict(fitLocalReg50,newdata =data.frame(age=ageTest),se=T)
lines(ageTest,predictLoess$fit,lwd=2,col="red")
lines(ageTest,predictLoess50$fit,lwd=2,col="blue")
legend("topright",legend = c("Span=0.2","Span=0.5"),col = c("red","blue"),cex=.8,lty=1,lwd=2)
```

Hooo, Welcome to GAM's Naresh
```{r}
gam1=lm(wage~ns(year,4)+ns(age,5)+education,data = Wage)
```
fitting gam's using gam library
```{r}
library(gam)
gamWage3=gam(wage~s(year,4)+s(age,5)+education,data = Wage)
par(mfrow=c(1,3))
plot.gam(gamWage3,se = T,col="red")
```
it is evident that year is linear in nature from above plots
test  with anova
```{r}
gamWage1=gam(wage~s(age,5)+education,data = Wage)
gamWage2=gam(wage~year+s(age,5)+education,data = Wage)
anova(gamWage1,gamWage2,gamWage3,test = "F")
summary(gamWage3)
plot.gam(gamWage2,se=T,col="red")
```
fitting local regression in gam's
```{r}
gamLocal=gam(wage~lo(year,age,span = 0.3)+education,data = Wage)
summary(gamLocal)
library(akima)
plot.gam(gamLocal,se=T,col="red")
```
fitting logistic regression with gam's
```{r}
gamLogitWage3=gam(I(wage>250)~s(year,4)+s(age,5)+education,data = Wage)
par(mfrow=c(1,3))
plot.gam(gamLogitWage3,se=T,col="blue")
table(education,I(wage>250))
gamLogitWage3=gam(I(wage>250)~s(year,4)+s(age,5)+education,data = Wage,subset = (education!="< HS Grad"))
par(mfrow=c(1,3))
plot.gam(gamLogitWage3,se=T,col="blue")
```
