---
title: "SVM_ISLR"
author: "NareshVemula"
date: "September 17, 2016"
output: html_document
---
bothe e1071 and liblinear libraries can be used to implement wide range of nonlinear algorithms
these can be used to implement SVM as well
we can use kernals such as linear , radial, polynomial
we will set the cost parameter which helps in selecting the support vectors to classify with errors less than cost given.
creating a matrix of observations to classify
```{r}
library(e1071)
set.seed(1)
x=matrix(rnorm(20*2),ncol=2)
y=c(rep(-1,10),rep(1,10))
str(x)
x[y==1,]=x[y==1,]+1
plot(x,col=(3-y))
```
changing y as factor variable
```{r}
svmDat=data.frame(x=x,y=as.factor(y))
svmFit=svm(y~.,data=svmDat,cost=10,kernal="linear",scale=FALSE)
plot(svmFit,svmDat)
svmFit$index
summary(svmFit)
```
using a smaller value for cost parameter
we can observe number of support vectors incresing as cost comes down
because the margine is wider
```{r}
svmFitCost=svm(y~.,data=svmDat,cost=.1,kernal="linear",scale=FALSE)
plot(svmFitCost,svmDat)
svmFitCost$index
summary(svmFitCost)
```
performing cv using tune parameter from e1071
```{r}
set.seed(1)
costCV=c(0.001 , 0.01, 0.1, 1,5,10,100)
svmCV=tune(svm,y~.,data = svmDat,kernel="linear",ranges=list(cost=costCV))
summary(svmCV)
bestModel=svmCV$best.model
summary(bestModel)
plot(bestModel,svmDat)
```
predicting using generated data
Support vector classifier
```{r}
xtest=matrix(rnorm(20*2),ncol=2)
ytest=sample(c(-1,1),20,rep=TRUE)
xtest[ytest ==1 ,]= xtest[ytest ==1,] + 1
testData=data.frame(x=xtest,y=as.factor(ytest))
predictTest=predict(bestModel,testData)
table(testData$y,predictTest)
```
application of svm on gene expression data
```{r}
library(ISLR)
data("Khan")
names(Khan)
dim(Khan$xtrain)
length(Khan$ytrain)
dim(Khan$xtest)
length(Khan$ytest)
table(Khan$ytest)
```
preparing data for svm
```{r}
geneTrain=data.frame(x=Khan$xtrain,y=as.factor(Khan$ytrain))
str(geneTrain)
geneTest=data.frame(x=Khan$xtest,y=as.factor(Khan$ytest))
geneCostCV=c(0.001 , 0.01, 0.1, 1,5,10,100)
```
cross validating 
```{r}
geneCV=tune(svm,y~.,data=geneTrain,kernel="linear",ranges=list(cost=geneCostCV))
summary(geneCV)
bestGeneModel=geneCV$best.model
summary(bestGeneModel)
table(bestGeneModel$fitted,geneTrain$y)
```
predicting on test data
```{r}
predictTest=predict(bestGeneModel,newdata=geneTest)
table(geneTest$y,predictTest)
```

