---
title: "Unsupervised Learning"
author: "NareshVemula"
date: "September 18, 2016"
output: html_document
---
Principal Component Analysis using US Arrests Data
--------------------------------------------------
```{r}
dimnames(USArrests)
str(USArrests)
apply(USArrests, 2, var)
apply(USArrests, 2, mean)
```
standardizing the variables having mean 0 and sd 1
```{r}
arrestPCA=prcomp(USArrests,scale=TRUE)
summary(arrestPCA)
arrestPCA
```
using the biplot we can observe the patterns where we have maximum crime rates and population from the USArrests Data
```{r}
biplot(arrestPCA,cex=.5,scale = 0,main="PCA of USArrests")
```
cluster analysis using generated data
```{r}
set.seed(101)
x=matrix(rnorm(100*2),ncol = 2,nrow = 100)
addMean=matrix(rnorm(8,sd = 4),4,2)
index=sample(1:4,100,replace = TRUE)
x=x+addMean[index,]
plot(x,col=index,pch=20,cex=2)
```
performnig kmeans clustering
============================
```{r}
generatedKmeans=kmeans(x,centers = 4,nstart = 15)
generatedKmeans
plot(x,col=generatedKmeans$cluster,cex=2,pch=1,lwd=2,main="kmeans cluster analysis")
points(x,col=index,pch=20,cex=2)
points(x,col=c(4,3,2,1)[index],pch=20,cex=2)

```
Hierarchical Clustering
======================
```{r}
GeneratedComplteClust=hclust(dist(x),method = "complete")
plot(GeneratedComplteClust)
GeneratedComplteClustWard=hclust(dist(x),method = "ward.D")
plot(GeneratedComplteClustWard)
```
cutting the tree at 4 clusters
```{r}
finalHclust=cutree(GeneratedComplteClust,4)
table(finalHclust,index)
```
comparing kmeans with hierarchial clustering
```{r}
table(finalHclust,generatedKmeans$cluster)
```
plotting dendogram with labels from index
```{r}
plot(GeneratedComplteClust,labels = index,cex=0.8)
```